\framework: Morphology-based Entity and Relational Entity Extraction Framework for Arabic

### 2018-04-20 tex/merf/TALMajor/answer/Reviewer1.comments.tex
done spell checking

  \item \label{Review1.1} 
The authors chose to use formal notation to describe the 
objects manipulated by the software. 
This make things rigouros but the notations are sometimes heavy.

@@

\item \label{Review1.5} 
how did you compute the time for the implementation of rules 
with the other system (one of the tasks needed 2 month of work 
!! is it really 2 acutal months of work or is it the period of 
time during which the development has been done)
more should be said about the evaluation methodology.

@@

\item \label{Review1.5}
you say nothing about the availability of the system, 
is it distributed ? free ? what kind of licence ?

>>>>

  \item \label{Review1.1} 
The authors chose to use formal notation to describe the 
objects manipulated by the software. 
This make things rigorous but the notations are sometimes heavy.

@@

\item \label{Review1.5} 
how did you compute the time for the implementation of rules 
with the other system (one of the tasks needed 2 month of work 
!! is it really 2 actual months of work or is it the period of 
time during which the development has been done)
more should be said about the evaluation methodology.

@@

\item \label{Review1.5}
you say nothing about the availability of the system, 
is it distributed ? free ? what kind of license ?

### 2018-04-20 tex/merf/TALMajor/answer/Reviewer3.comments.tex
done spell checking

This work is an improvement step in a continuous previous work 
(Saref,MATAr) which leads to a complete project that can be 
used as a tool for other researchers in the field to help 
improving the Arabic language computations and processings.

It depends on the morphological analyser (Saref) results 
to define tags for flexible corpus annotation that is automatic 
but allow the users to correct and fix the tagging. 
In additions it extract entitiy and relational entityies from 
the input document.

@@

\item \label{Review.3.3}
2- ``How to use MERF?''
(you can have more explanations in your appendex and could have 
``help'' button in your API. This part was also relatively 
clear for the reader.

@@

\item \label{Review.3.11}
4- Section 5 can be combined with section 2. 
you can introduce the user friendly interface using the 
specified example. 
Section 2 seemed not very imporatnt as stand alone section.

@@

\answer{
  We added  the following to the discussion Section. 
  \begin{quote}
We notice that ANGE, ATEEMA, and Genealogy tree report higher 
    precision than \framework. 
This is mainly due to their capacity 
to learn words and relations that may not have a match in the 
morphological analyzer based on co-occurence relations. 
For example, the sequence $p_1 t_1 p_2$ where $p_1$ and 
$p_2$ are persons and $t_1$ is a tell relationship helps
indicate that $x$ is a tell relationship in $p_1 x p_2$ 
even if the morphological analyzer did not return the required
feature for $x$ to match a tell relationship. 
\framework does not have that capacity yet unless it is
encoded in the C++ actions. 
  \end{quote} 
}

\item \label{Review.3.28}
Clarity:  
It was not very clear how the work was done 
``Methodology part''. 
Section 4 ``MERF'' please add ``Methedology''.

@@

\answer{
  We clarify the 
  notation ``up to $M$'' in the paper to be as follows. 
  \begin{quote}
   up to $M$ repititions where $M$ is 
   a non-zero positive integer 
  \end{quote}
}

@@

\answer{
  We used italics to emphasize important 
  concepts that we introduce for the first time. 
  We reduced the use of emphasid text in the paper. 
}

>>>>

This work is an improvement step in a continuous previous work 
(Saref,MATAr) which leads to a complete project that can be 
used as a tool for other researchers in the field to help 
improving the Arabic language computations and processing.

It depends on the morphological analyser (Saref) results 
to define tags for flexible corpus annotation that is automatic 
but allow the users to correct and fix the tagging. 
In additions it extract entity and relational entities from 
the input document.

@@

\item \label{Review.3.3}
2- ``How to use MERF?''
(you can have more explanations in your appendix and could have 
``help'' button in your API. This part was also relatively 
clear for the reader.

@@

\item \label{Review.3.11}
4- Section 5 can be combined with section 2. 
you can introduce the user friendly interface using the 
specified example. 
Section 2 seemed not very important as stand alone section.

@@

\answer{
  We added  the following to the discussion Section. 
  \begin{quote}
We notice that ANGE, ATEEMA, and Genealogy tree report higher 
    precision than \framework. 
This is mainly due to their capacity 
to learn words and relations that may not have a match in the 
morphological analyzer based on co-occurrence relations. 
For example, the sequence $p_1 t_1 p_2$ where $p_1$ and 
$p_2$ are persons and $t_1$ is a tell relationship helps
indicate that $x$ is a tell relationship in $p_1 x p_2$ 
even if the morphological analyzer did not return the required
feature for $x$ to match a tell relationship. 
\framework does not have that capacity yet unless it is
encoded in the C++ actions. 
  \end{quote} 
}

\item \label{Review.3.28}
Clarity:  
It was not very clear how the work was done 
``Methodology part''. 
Section 4 ``MERF'' please add ``Methodology''.

@@

\answer{
  We clarify the 
  notation ``up to $M$'' in the paper to be as follows. 
  \begin{quote}
   up to $M$ repetitions where $M$ is 
   a non-zero positive integer 
  \end{quote}
}

@@

\answer{
  We used italics to emphasize important 
  concepts that we introduce for the first time. 
  We reduced the use of emphasized text in the paper. 
}

### 2018-04-20 tex/merf/TALMajor/discussion.tex
done spell checking

We notice that ANGE, ATEEMA, and Genealogy tree report higher precision than \framework. 
This is mainly due to their capacity 
to learn words and relations that may not have a match in the 
morphological analyzer based on co-occurence relations. 
For example, the sequence $p_1 t_1 p_2$ where $p_1$ and 
$p_2$ are persons and $t_1$ is a tell relationship helps
indicate that $x$ is a tell relationship in $p_1 x p_2$ 
even if the morphological analyzer did not return the required
feature for $x$ to match a tell relationship. 
\framework does not have that capacity yet unless it is
encoded in the C++ actions. 
}

>>>>

We notice that ANGE, ATEEMA, and Genealogy tree report higher precision than \framework. 
This is mainly due to their capacity 
to learn words and relations that may not have a match in the 
morphological analyzer based on co-occurrence relations. 
For example, the sequence $p_1 t_1 p_2$ where $p_1$ and 
$p_2$ are persons and $t_1$ is a tell relationship helps
indicate that $x$ is a tell relationship in $p_1 x p_2$ 
even if the morphological analyzer did not return the required
feature for $x$ to match a tell relationship. 
\framework does not have that capacity yet unless it is
encoded in the C++ actions. 
}

### 2018-04-20 tex/merf/TALMajor/introduction.tex
done spell checking

In this paper, 
we present \framework, a morphology-based entity and relational entity
extraction framework for Arabic text.
\framework provides a user-friendly interface where the user defines tag types 
and associates them with
%\framework formulae that are 
regular expressions over 
%\framework
Boolean formulae.
%
A Boolean formula is defined by a term, negation of a term, or disjunction of terms.
Terms are matches to Arabic morphological features including 
prefix, stem, suffix, part of speech (POS) tags, gloss tags, extended synonym 
tags, and semantic categories.
\major{
For example, entity $p_1$ in Figure~\ref{fig:intromotiv} has a place semantic 
category. 
}
\framework regular expressions support operators such as concatenation, 
zero or one, zero or more, one or more, up to $M$ \major{repititions where $M$ is 
a non-zero positive integer}, and logical conjunction and disjunction.
\major{
For example, the sequence between $p_1$ and $p_2$ matches a regular expression $re$ 
that requires two semantic place categories with a place-preposition POS 
tag ($r_1$) in between. 
}

@@

In \framework, we make the following contributions.
\begin{itemize}
  \item \framework enables the user to define relations in a simple manner
    and automatically detects relational entities matching the user defined relations. 
  \item \framework enables the user to associate sub-expressions
    with code actions, and executes the code action 
    when a corresponding match is found.
    It also provides an API to enable access to match 
    features such as text, position, length, numerical value, and morphological features.
  \item \framework enables the user to tag words based on a 
    \major{
    novel ligh Arabic wordnet relation that leverages the 
    synonym $Syn^k$ feature.
  }

>>>>

In this paper, 
we present \framework, a morphology-based entity and relational entity
extraction framework for Arabic text.
\framework provides a user-friendly interface where the user defines tag types 
and associates them with
%\framework formulae that are 
regular expressions over 
%\framework
Boolean formulae.
%
A Boolean formula is defined by a term, negation of a term, or disjunction of terms.
Terms are matches to Arabic morphological features including 
prefix, stem, suffix, part of speech (POS) tags, gloss tags, extended synonym 
tags, and semantic categories.
\major{
For example, entity $p_1$ in Figure~\ref{fig:intromotiv} has a place semantic 
category. 
}
\framework regular expressions support operators such as concatenation, 
zero or one, zero or more, one or more, up to $M$ \major{repetitions where $M$ is 
a non-zero positive integer}, and logical conjunction and disjunction.
\major{
For example, the sequence between $p_1$ and $p_2$ matches a regular expression $re$ 
that requires two semantic place categories with a place-preposition POS 
tag ($r_1$) in between. 
}

@@

In \framework, we make the following contributions.
\begin{itemize}
  \item \framework enables the user to define relations in a simple manner
    and automatically detects relational entities matching the user defined relations. 
  \item \framework enables the user to associate sub-expressions
    with code actions, and executes the code action 
    when a corresponding match is found.
    It also provides an API to enable access to match 
    features such as text, position, length, numerical value, and morphological features.
  \item \framework enables the user to tag words based on a 
    \major{
    novel light Arabic wordnet relation that leverages the 
    synonym $Syn^k$ feature.
  }

### 2018-04-20 tex/merf/TALMajor/relatedwork.tex
done spell checking

\begin{table}[tb!]
\caption{Comaprison of \framework with SystemT, 
TEXTMARKER, 
Urbain, 
QARAB}
\resizebox{\columnwidth}{!}{
\begin{tabular}{l|c|c|c|c|c}
%{p{2.7cm}|p{.7cm}|p{1.2cm}|p{1.2cm}|p{1.2cm}|p{1.2cm}}
Features & \framework & SystemT & TEXT\linebreak MARKER & Urbain & QARAB\\ \hline
Query type & MRE & AQL & matching \linebreak rules & natural language & natural language \\ \hline
Morphology support& $\checked$ & - & - & OpenNLP & Parser \\ \hline
Relations & $\checked$ & - & - & $\checked$ & - \\ \hline
Actions & $\checked$ & - & - & - & - \\ \hline
Editor & $\checked$ & - & $\checked$ & - & - \\ \hline
Tag visualization & $\checked$ & - & $\checked$ & - & - \\ \hline
Graph visualization & $\checked$ & - & - & - & - \\
\end{tabular}
}
\label{tab:iecomp}
\end{table}

@@

\major{
Another track in the literature targets specific tasks such as NER using 
statistical and machine learning techniques such as maximum entropy, 
optimized feature sets and conditional random fields~\cite{benajiba2007anersys,benajiba2008arabic,ekbal2010named,abdelrahman2010integrated}.
}
%The work in~\cite{ekbal2010named} presents a language independent approach for NER extraction using {\em support vector machines}. 
%The work in~\cite{abdelrahman2010integrated} integrates a semi-supervised bootstrapping pattern recognition technique, and a supervised classifier based on {\em conditional random fields} to solve NER problems.
Knowledge-based techniques such 
as~\cite{zaghouani2010adapting,traboulsi2009arabic} propose local grammars with morphological stemming.% to perform NER. 
%The work in~
\cite{ZaMaHaCicling2012Entity} present a method for extracting entities and events, and relations amongst them from Arabic text using a hierarchy of manually built 
finite state machines driven by morphological features, 
and graph transformation algorithms. 
Such techniques require advanced linguistic and programming expertise.

@@

{\bf Annotation tools.~}
MMAX2 is a manual multi-level linguistic annotation tool with an XML 
based data model~\cite{mmax2}.
%It enables the user to create, browse, visualize, and query annotations and may be able to resolve coreference tags. 
BRAT~\cite{brat} and WordFreak~\cite{wordfreak} are manual 
multi-lingual user friendly web-based annotators that allow the construction 
of entity and relation annotation corpora.
%They can be extended through plug-ins to enable automatic annotators and customized annotation visualization specifications.
Knowtator~\cite{ogren2006knowtator} is a general purpose incremental text annotation tool 
implemented as a Prot\'eg\'e~\cite{gennari2003evolution} plug-in. 
Prot\'eg\'e is an open-source platform with a suite of tools to construct domain models and knowledge-based applications with ontologies. 
However, it doesn't support the Arabic language.

@@

Knowledge-based techniques such 
as~\cite{zaghouani2010adapting,traboulsi2009arabic} propose local grammars 
with morphological stemming to perform NER. 
%The work in~
\cite{ZaMaHaCicling2012Entity} presents a method for extracting entities, 
events, and relations amongst them from Arabic text using a hierarchy of manually built 
finite state machines driven by morphological features and graph 
transformation algorithms. 
Such techniques require advanced linguistic and programming expertise.

>>>>

\begin{table}[tb!]
\caption{Comparison of \framework with SystemT, 
TEXTMARKER, 
Urbain, 
QARAB}
\resizebox{\columnwidth}{!}{
\begin{tabular}{l|c|c|c|c|c}
%{p{2.7cm}|p{.7cm}|p{1.2cm}|p{1.2cm}|p{1.2cm}|p{1.2cm}}
Features & \framework & SystemT & TEXT\linebreak MARKER & Urbain & QARAB\\ \hline
Query type & MRE & AQL & matching \linebreak rules & natural language & natural language \\ \hline
Morphology support& $\checked$ & - & - & OpenNLP & Parser \\ \hline
Relations & $\checked$ & - & - & $\checked$ & - \\ \hline
Actions & $\checked$ & - & - & - & - \\ \hline
Editor & $\checked$ & - & $\checked$ & - & - \\ \hline
Tag visualization & $\checked$ & - & $\checked$ & - & - \\ \hline
Graph visualization & $\checked$ & - & - & - & - \\
\end{tabular}
}
\label{tab:iecomp}
\end{table}

@@

\major{
Another track in the literature targets specific tasks such as NER using 
statistical and machine learning techniques such as maximum entropy, 
optimized feature sets and conditional random fields~\cite{benajiba2007anersys,benajiba2008arabic,ekbal2010named,abdelrahman2010integrated}.
}
%The work in~\cite{ekbal2010named} presents a language independent approach for NER extraction using {\em support vector machines}. 
%The work in~\cite{abdelrahman2010integrated} integrates a semi-supervised bootstrapping pattern recognition technique, and a supervised classifier based on {\em conditional random fields} to solve NER problems.
Knowledge-based techniques such 
as~\cite{zaghouani2010adapting,traboulsi2009arabic} propose local grammars with morphological stemming.% to perform NER. 
%The work in~
\cite{ZaMaHaCicling2012Entity} present a method for extracting entities and events, and relations among them from Arabic text using a hierarchy of manually built 
finite state machines driven by morphological features, 
and graph transformation algorithms. 
Such techniques require advanced linguistic and programming expertise.

@@

{\bf Annotation tools.~}
MMAX2 is a manual multi-level linguistic annotation tool with an XML 
based data model~\cite{mmax2}.
%It enables the user to create, browse, visualize, and query annotations and may be able to resolve coreference tags. 
BRAT~\cite{brat} and WordFreak~\cite{wordfreak} are manual 
multi-lingual user friendly web-based annotators that allow the construction 
of entity and relation annotation corpora.
%They can be extended through plug-ins to enable automatic annotators and customized annotation visualization specifications.
Knowtator~\cite{ogren2006knowtator} is a general purpose incremental text annotation tool 
implemented as a Prot\'eg\'e~\cite{gennari2003evolution} plug-in. 
Prot\'eg\'e is an open-source platform with a suite of tools to construct domain models and knowledge-based applications with ontology. 
However, it doesn't support the Arabic language.

@@

Knowledge-based techniques such 
as~\cite{zaghouani2010adapting,traboulsi2009arabic} propose local grammars 
with morphological stemming to perform NER. 
%The work in~
\cite{ZaMaHaCicling2012Entity} presents a method for extracting entities, 
events, and relations among them from Arabic text using a hierarchy of manually built 
finite state machines driven by morphological features and graph 
transformation algorithms. 
Such techniques require advanced linguistic and programming expertise.

### 2018-04-20 tex/merf/TALMajor/results.tex
done spell checking

We evaluated ANGE with
Musnad Ahmad, a hadith book, where we constructed an annotated 
golden reference containing 1,865 words.
We evaluated ATEEMA 
with articles from issues of  the Lebanese 
Al-Akhbar newspaper where we constructed 
an annotated golden reference containing 1,677 words. 
For the genealogical tree extraction we used 
an extract from the Genesis biblical text with 1,227 words.
Finally, we used an annotated article from the Lebanese Assafir
newspaper with 1,399 words to evaluate the NUMNORM case study
\footnote{available at \url{http://www.assafir.com} and 
\url{http://www.al-akhbar.com}.}. 
%
In the online appendix%
~\footnote{available at ~\url{-ommited-for-anonymity-}}%http://webfea.fea.aub.edu.lb/fadi/pdfs/merfappendix.pdf
, we report on eight additional \framework case studies.
\major{
Manual annotators inspected the outcome and provided corrections where tools 
made mistakes.
The corrections form the manual gold annotation that we compared against.
}

@@

The first row in Table~\ref{tab:nchain} is an example narrator chain,
the second is the transliteration, the third 
shows the MBF tags. Rows 4, 5, and 6 show the 
matches for \cci{name}, \cci{nar}, and \cci{nchain},
respectively.
%
\framework assigns the symbols $s_1$ and $s_2$ for the 
MRE sub-expressions \cci{TOLD} and \cci{nar}, respectively. 
We define the relation $\langle s_2,s_2',s_1\rangle$ 
to relate sequences of narrators with edges labelled by the tags of \cci{TOLD} where 
$s_2'$ denotes the next match of \cci{nar} in the one or more MRE subexpression.
%The narrators in the example shown in Table~\ref{tab:nchain} are \RL{qtybT bn s`yd}, \RL{jryr}, and \RL{`mArT bn Alq`qA`}. 
%\transfalse
%The edges relating the entities are labeled by the word set \{\RL{.hdd_tnA}, \RL{.hdd_tnA}, \RL{`n}\} which contains all the %matches of the MBF \cci{TOLD}.
%\transtrue
%
Table~\ref{tab:mbfer} shows that \framework detected almost all the MBF matches 
with 99\% recall and 85\% precision and 
extracted user-defined relations with 98\% recall and 99\% precision.

>>>>

We evaluated ANGE with
Musnad Ahmad, a hadith book, where we constructed an annotated 
golden reference containing 1,865 words.
We evaluated ATEEMA 
with articles from issues of  the Lebanese 
Al-Akhbar newspaper where we constructed 
an annotated golden reference containing 1,677 words. 
For the genealogical tree extraction we used 
an extract from the Genesis biblical text with 1,227 words.
Finally, we used an annotated article from the Lebanese Assafir
newspaper with 1,399 words to evaluate the NUMNORM case study
\footnote{available at \url{http://www.assafir.com} and 
\url{http://www.al-akhbar.com}.}. 
%
In the online appendix%
~\footnote{available at ~\url{-omitted-for-anonymity-}}%http://webfea.fea.aub.edu.lb/fadi/pdfs/merfappendix.pdf
, we report on eight additional \framework case studies.
\major{
Manual annotators inspected the outcome and provided corrections where tools 
made mistakes.
The corrections form the manual gold annotation that we compared against.
}

@@

The first row in Table~\ref{tab:nchain} is an example narrator chain,
the second is the transliteration, the third 
shows the MBF tags. Rows 4, 5, and 6 show the 
matches for \cci{name}, \cci{nar}, and \cci{nchain},
respectively.
%
\framework assigns the symbols $s_1$ and $s_2$ for the 
MRE sub-expressions \cci{TOLD} and \cci{nar}, respectively. 
We define the relation $\langle s_2,s_2',s_1\rangle$ 
to relate sequences of narrators with edges labeled by the tags of \cci{TOLD} where 
$s_2'$ denotes the next match of \cci{nar} in the one or more MRE subexpression.
%The narrators in the example shown in Table~\ref{tab:nchain} are \RL{qtybT bn s`yd}, \RL{jryr}, and \RL{`mArT bn Alq`qA`}. 
%\transfalse
%The edges relating the entities are labeled by the word set \{\RL{.hdd_tnA}, \RL{.hdd_tnA}, \RL{`n}\} which contains all the %matches of the MBF \cci{TOLD}.
%\transtrue
%
Table~\ref{tab:mbfer} shows that \framework detected almost all the MBF matches 
with 99\% recall and 85\% precision and 
extracted user-defined relations with 98\% recall and 99\% precision.
